Tropical Andes plant traits

### 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd

# Load required packages
```{r}
library(dplyr)
library(tidyr)
```

@@

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
#write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

```{r}
#need to remove none & NA
plant_traits_dispersal <- plant_traits_dispersal[plant_traits_dispersal$trait_value == "biotic" | plant_traits_dispersal$trait_value == "abiotic",]
#need one value per species

plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
```

@@

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
#trait values are a mix of numeric and characters
```

# Combine all traits
```{r}
# combine all trait wide tables
plant_traits_wide <- bind_cols(plant_traits_height_wide, plant_traits_mass_wide)
# remove species with NAs for one or more traits

```

>>>>

# Load required packages
```{r}
library(dplyr)
library(tidyr)
library(data.table)
```

@@

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

```{r}
# change all trait_name to dispersal_syndrome
plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'Dispersal syndrome'] <- 'dispersal_syndrome'
plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'whole plant dispersal syndrome'] <- 'dispersal_syndrome'
#need to remove none & NA
plant_traits_dispersal <- plant_traits_dispersal[plant_traits_dispersal$trait_value == "biotic" | plant_traits_dispersal$trait_value == "abiotic",]
#need one value per species
plant_traits_dispersal <- plant_traits_dispersal %>% distinct(species, .keep_all = TRUE)
plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
#ready to combine
```

@@

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
# change all trait_name to plant lifespan
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'Plant lifespan (longevity)'] <- 'plant_lifespan'
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'maximum whole plant longevity'] <- 'plant_lifespan'

#remove non-numeric values
plant_traits_longevity <- plant_traits_longevity[!is.na(as.numeric(as.character(plant_traits_longevity$trait_value))),]
```

```{r}
plant_traits_longevity$trait_value <- as.numeric(plant_traits_longevity$trait_value)
plant_traits_longevity_wide <- pivot_wider(data = plant_traits_longevity, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
```

# Combine all traits
```{r}
# convert all to dataframes
plant_traits_height_wide <- as.data.frame(plant_traits_height_wide)
plant_traits_dispersal_wide <- as.data.frame(plant_traits_dispersal_wide)
plant_traits_mass_wide <- as.data.frame(plant_traits_mass_wide)
plant_traits_longevity_wide <- as.data.frame(plant_traits_longevity_wide)

plant_traits_wide <- merge(x = plant_traits_height_wide, y = plant_traits_dispersal_wide, by = "species", all = TRUE)
plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_mass_wide, by = "species", all = TRUE)
plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_longevity_wide, by = "species", all = TRUE)
```

```{r}
plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "Fruit dry mass", "plant_lifespan")]
# remove species with NAs for one or more traits
plant_traits_wide_complete <- na.omit(plant_traits_wide)
```

#Summary
```{r}
glimpse(plant_traits_wide_complete)
```
